\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-02-23 papers/methods/paper.tex
Adding paper conclusion. Tongue in cheek ending. Lets see if it passes review (ha)

Points to make:

1. Created more efficient BHM model with treatment of selection effects, populations, unexplained dispersion, systematics.
2. Main issue is still scatter model. Need to characterise this theoretically because adding full variance for them into the model results in weak cosmology.
3. Bias in observed summary statistics is difficult to model analytically, and future work should incorporate bias corrections calculated from simulations. However, even numerically calculated bias corrections run into issues in that the type Ia dispersion is not well characterised.
4. Model will be applied shortly to the DES 3yr spectroscopic sample.

>>>>

In this paper we have outlined the creation of a Hierarchical Bayesian model for supernova cosmology. The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter colour smearing and incorporates uncertainty from peculiar velocity uncertainty, calibration uncertainty, dust uncertainty and SALT2 model uncertainty. The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters and run in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to quadratic or cubic run times for models that rely on matrix inversion.

The importance of validating models using high-precision statistics gained by performing fits to hundreds of data realisations cannot be overstated, however is lacking in many prior BHM models for supernova cosmology. We have validated this model against many realisations of simplistic simulations with well-known and well-defined statistics, and found no bias. When validating using SNANA simulations, we find evidence of bias which is traced back to light curve fits reporting biased observables and incorrect covariance. Allowing fully parametrised corrections on observed supernovae summary statistics is found to make cosmology fits too weak, and allowing simulation based corrections to vary in strength is found to give minor reductions in bias, however the uncertainty on the intrinsic scatter model itself limits the efficacy of the bias corrections. For the data size represented in the DES three-year spectroscopic survey, the determined biases should be sub-dominant to other sources of uncertainty, however this cannot be expected for future analyses with larger datasets. Stricter bias corrections calculated from simulations are required to reduce bias. Ideally, this would include further work on the physical intrinsic dispersion of the Type Ia supernovae population such that we can better characterise this bias.

With our model being validated against hundreds of simulation realisations, representing a combined dataset over more than $250\,000$ simulated supernovae, we have been able to accurately determine biases in our model and trace their origin. With the current biases being sub-dominant to the total uncertainty, we now prepare to analyse the DES three-ear dataset. But that, my friends, is another, much more important, paper.
